On the spatial aggregation of ecosystem condition indicators

### 2025-01-08 manuscript/manuscript.qmd
Change to authoryear

keywords: 
  - SEEA EA
  - ecosystem condition
  - ecosystem accounting
  - indicators
  - aggregation bias
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: ---
      formatting: preprint
      model: 3p
      cite-style: super
    include-in-header:
      text: '\usepackage{lineno}\linenumbers'
execute:
  cache: true
  warning: false
  message: false
editor: 
  markdown: 
    wrap: sentence
---

@@

*Conclusion*: Because of how X~100~ is defined based on the median value across reference lakes, overshooting values (\> X~100~) should be preserved in the spatial aggregation.
<!-- Otherwise, the condition of water bodies with X > X80 will be underreported.  --> Otherwise we get a negative displacement, as in the bilberry example.
This means that neither EQR values (which are untransformed), nor nEQR values (which are truncated), can be used uncritically in ECAs.
One solution, to enable the use of WFD indicators in ECA, would be to change to pathway 6, with (1) scaling, (2) transforming, (3) aggregating and (4) truncating the values [@fig.pathways].
Furthermore, Eurostat @eurostat2022\
<!-- requires  --> encourages reporting of variable values.
These, however, cannot be aggregated spatially in a meaningful way.
<!-- because the use of threshold values amounts to a nonlinear (specifically, a piecewise-linear) transformation.  -->

>>>>

keywords: 
  - SEEA EA
  - ecosystem condition
  - ecosystem accounting
  - indicators
  - aggregation bias
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: ---
      formatting: preprint
      model: 3p
      cite-style: authoryear
    include-in-header:
      text: '\usepackage{lineno}\linenumbers'
execute:
  cache: true
  warning: false
  message: false
editor: 
  markdown: 
    wrap: sentence
---

@@

*Conclusion*: Because of how X~100~ is defined based on the median value across reference lakes, overshooting values (\> X~100~) should be preserved in the spatial aggregation.
<!-- Otherwise, the condition of water bodies with X > X80 will be underreported.  --> Otherwise we get a negative displacement, as in the bilberry example.
This means that neither EQR values (which are untransformed), nor nEQR values (which are truncated), can be used uncritically in ECAs.
One solution, to enable the use of WFD indicators in ECA, would be to change to pathway 6, with (1) scaling, (2) transforming, (3) aggregating and (4) truncating the values [@fig-pathways].
Furthermore, Eurostat @eurostat2022\
<!-- requires  --> encourages reporting of variable values.
These, however, cannot be aggregated spatially in a meaningful way.
<!-- because the use of threshold values amounts to a nonlinear (specifically, a piecewise-linear) transformation.  -->
